---
title: "Reverse Engineering Project"
author: "Student names here"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [STORY NAME](STORY URL)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data
baltimore_homicide <- read_csv("data/homicide-data.csv")
# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 


```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: "City police made an arrest in 41 percent of homicides in 2014; last year, the rate was just 27 percent, a 14 percentage point drop."
* **Analysis summary**: We were able to confirm the findings in this sentence, 2014 had 41% and 2017 had 27%. 2014 we founded up .2 and 2017 we rounded down .3 but the finding in the story was not wrong according to our data analysis.

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock
percent_2014<-baltimore_homicide %>%
filter(city == "Baltimore") %>%
mutate(reported_date = as.Date(as.character( reported_date), format="%Y%m%d")) %>%
filter(reported_date >= as.Date("2014-01-01"), reported_date <= as.Date("2014-12-31")) %>%
group_by(disposition)%>%
summarise(
count= n()
) %>%
mutate(total_homicides = sum(count)) %>%
filter(disposition == "Closed by arrest") %>%
mutate(c = sum(count)) %>%
mutate(percentage = count / total_homicides) 

#2017
percent_2017<-baltimore_homicide %>%
filter(city == "Baltimore") %>%
mutate(reported_date = as.Date(as.character( reported_date), format="%Y%m%d")) %>%
filter(reported_date >= as.Date("2017-01-01"), reported_date <= as.Date("2017-12-31")) %>%
group_by(disposition)%>%
summarise(
count= n()
) %>%
mutate(total_homicides = sum(count)) %>%
filter(disposition == "Closed by arrest") %>%
mutate(count = sum(count)) %>%
mutate(percentage = count / total_homicides) 

total_2014_2017<-bind_cols(percent_2014,percent_2017)
```

### Sentence 2

* **Sentence text**: In Chicago, the homicide arrest rate has dropped 21 percentage points, in Boston it has dropped 12 points and in St. Louis it is down 9. (2014 vs 2017)
* **Analysis summary**: Through filtering the data to only homicides that resulted in an arrest in 2014 and 2015 in Chicago, we subtracted the percent of total homicides from each year to confirm that the homicide arrest rate in Chicago has dropped 21 percentage points. We did this same process for the cities of Boston and St. Louis to confirm that Boston dropped 12 percentage points and St. Louis dropped 9 percentage points.

```{r}
# Put code to reverse engineer sentence here
#Chicago 2014
chicago_percent_2014<- baltimore_homicide %>% 
  filter(city == "Chicago") %>%
  mutate(reported_date = as.Date(as.character( reported_date), format="%Y%m%d")) %>%
  filter(reported_date >= as.Date("2014-01-01"), reported_date <= as.Date("2014-12-31")) %>%
group_by(disposition)%>%
   summarise(
    count= n()
  ) %>%
  mutate(total_homicides = sum(count)) %>%
  filter(disposition == "Closed by arrest") %>%
  mutate(percentagechicago2014 = count / total_homicides) 
  
#Chicago 2017
chicago_percent_2017<- baltimore_homicide %>% 
  filter(city == "Chicago") %>%
  mutate(reported_date = as.Date(as.character( reported_date), format="%Y%m%d")) %>%
  filter(reported_date >= as.Date("2017-01-01"), reported_date <= as.Date("2017-12-31")) %>%
group_by(disposition)%>%
   summarise(
    count= n()
  ) %>%
  mutate(total_homicides = sum(count)) %>%
  filter(disposition == "Closed by arrest") %>%
  mutate(percentagechicago2017 = count / total_homicides) 
#CHICAGO PERCENT CHANGE
chicago_total <- bind_cols(chicago_percent_2014,chicago_percent_2017)%>%
  mutate(percentagechicago2017-percentagechicago2014)
#Boston
#Boston 2014
boston_2014 <- baltimore_homicide %>% 
  filter(city == "Boston") %>%
  mutate(reported_date = as.Date(as.character( reported_date), format="%Y%m%d")) %>%
  filter(reported_date >= as.Date("2014-01-01"), reported_date <= as.Date("2014-12-31")) %>%
group_by(disposition)%>%
   summarise(
    count= n()
  ) %>%
  mutate(total_homicides = sum(count)) %>%
  filter(disposition == "Closed by arrest") %>%
  mutate(percentageboston2014 = count / total_homicides)
#Boston 2017
boston_2017<- baltimore_homicide %>% 
  filter(city == "Boston") %>%
  mutate(reported_date = as.Date(as.character( reported_date), format="%Y%m%d")) %>%
  filter(reported_date >= as.Date("2017-01-01"), reported_date <= as.Date("2017-12-31")) %>%
group_by(disposition)%>%
   summarise(
    count= n()
  ) %>%
  mutate(total_homicides = sum(count)) %>%
  filter(disposition == "Closed by arrest") %>%
  mutate(percentageboston2017 = count / total_homicides) 
  
#Boston Percent Change
boston_total <- bind_cols(boston_2014,boston_2017)%>%
  mutate(percentageboston2017-percentageboston2014)
  
#St. Louis
#St. Louis 2014
St_Louis_2014 <- baltimore_homicide %>% 
  filter(city == "St. Louis") %>%
  mutate(reported_date = as.Date(as.character( reported_date), format="%Y%m%d")) %>%
 filter(reported_date >= as.Date("2014-01-01"), reported_date <= as.Date("2014-12-31")) %>%
group_by(disposition)%>%
   summarise(
    count= n()
  ) %>%
  mutate(total_homicides = sum(count)) %>%
  filter(disposition == "Closed by arrest") %>%
  mutate(percentage_stlouis_2014 = count / total_homicides)
 
 
#St. Louis 2017
St_Louis_2017 <- baltimore_homicide %>% 
  filter(city == "St. Louis") %>%
  mutate(reported_date = as.Date(as.character( reported_date), format="%Y%m%d")) %>%
 filter(reported_date >= as.Date("2017-01-01"), reported_date <= as.Date("2017-12-31")) %>%
group_by(disposition)%>%
   summarise(
    count= n()
  ) %>%
  mutate(total_homicides = sum(count)) %>%
  filter(disposition == "Closed by arrest") %>%
  mutate(percentage_stlouis_2017 = count / total_homicides)
#St. Louis Percent Change
St_Louis_total <- bind_cols(St_Louis_2014,St_Louis_2017)%>%
  mutate(percentage_stlouis_2017-percentage_stlouis_2014)
  
# Display results of code below this codeblock

```

### Sentence 3

* **Sentence text**: Of the 1,002 homicides between 2015 and the beginning of this year, just 252 — one out of every four — resulted in an arrest.


* **Analysis summary**: Through filtering the data by homicides between the beginning of 2015 to the end of 2017, we can confirm that there were 1,002 homicides in this time period. By grouping the data by dispositions listed as "closed by arrest", we can confirm that 252 cases resulted in arrest. By finding the percent of cases resulting in arrest by the total homicide cases, we can confirm that one out of ever four case resulted in an arrest.

```{r}
# Put code to reverse engineer sentence here
baltimore_homicide %>% 
  filter(city == "Baltimore") %>%
  mutate(reported_date = as.Date(as.character( reported_date), format="%Y%m%d")) %>%
  filter(reported_date >= as.Date("2015-01-01"), reported_date <= as.Date("2017-12-31")) %>%
group_by(disposition)%>%
   summarise(
    count= n()
  ) %>%
  mutate(total_homicides = sum(count)) %>%
  filter(disposition == "Closed by arrest") %>%
  mutate(percentage = count / total_homicides) 
# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: Baltimore is also one of 30 cities that have seen an increase in homicides in recent years, with the greatest raw number increase in killings of any city other than Chicago, which has four times the population.
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here
#Code that organizes top 30 cities (2014 v 2017)
homicide_2014 <- baltimore_homicide %>%
  mutate(reported_date = as.Date(as.character( reported_date), 
  format="%Y%m%d")) %>%
  filter(reported_date >= as.Date("2014-01-01"), reported_date <= as.Date("2014-12-31")) %>%
  group_by(city) %>%
  summarise(
    count= n()) %>%
   arrange(desc(count))
homicide_2017 <- baltimore_homicide %>%
  mutate(reported_date = as.Date(as.character( reported_date), 
  format="%Y%m%d")) %>%
  filter(reported_date >= as.Date("2017-01-01"), reported_date <= as.Date("2017-12-31")) %>%
  group_by(city) %>%
  summarise(
    count= n()) %>%
   arrange(desc(count)) %>%
  rename(count2 = count)
alldata <- homicide_2014 %>%
  inner_join(homicide_2017)
alldata <- alldata %>%
  mutate(difference=count2 - count) %>%
  arrange(desc(difference))

# Display results of code below this codeblock

```

