---
title: "Reverse Engineering Project"
author: "Student names here"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [STORY NAME](STORY URL)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data
baltimore_homicide <- read_csv("data/homicide-data.csv")
# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 


```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: "City police made an arrest in 41 percent of homicides in 2014; last year, the rate was just 27 percent, a 14 percentage point drop."
* **Analysis summary**: We were able to confirm the findings in this sentence, 2014 had 41% and 2017 had 27%. 2014 we founded up .2 and 2017 we rounded down .3 but the finding in the story was not wrong according to our data analysis.

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock
percent_2014<-baltimore_homicide %>%
filter(city == "Baltimore") %>%
mutate(reported_date = as.Date(as.character( reported_date), format="%Y%m%d")) %>%
filter(reported_date >= as.Date("2014-01-01"), reported_date <= as.Date("2014-12-31")) %>%
group_by(disposition)%>%
summarise(
count= n()
) %>%
mutate(total_homicides = sum(count)) %>%
filter(disposition == "Closed by arrest") %>%
mutate(c = sum(count)) %>%
mutate(percentage = count / total_homicides) 

#2017
percent_2017<-baltimore_homicide %>%
filter(city == "Baltimore") %>%
mutate(reported_date = as.Date(as.character( reported_date), format="%Y%m%d")) %>%
filter(reported_date >= as.Date("2017-01-01"), reported_date <= as.Date("2017-12-31")) %>%
group_by(disposition)%>%
summarise(
count= n()
) %>%
mutate(total_homicides = sum(count)) %>%
filter(disposition == "Closed by arrest") %>%
mutate(count = sum(count)) %>%
mutate(percentage = count / total_homicides) 

```

### Sentence 2

* **Sentence text**: "Of 50 of the nation’s largest cities, Baltimore is one of 34 where police now make homicide arrests less often than in 2014."
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 3

* **Sentence text**: Of the 1,002 homicides between 2015 and the beginning of this year, just 252 — one out of every four — resulted in an arrest.


* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: Sandtown-Winchester, where Gray died, has seen 22 more homicides in the three-year period since Gray’s death than it did in the three years before he died. Southwest Baltimore saw its homicides rise by 35, and Greater Rosemont has seen 26 more since 2015.
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: In each of those neighborhoods, police make an arrest in fewer than 25 percent of cases, including 16 percent in Sandtown-Winchester.

* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-