---
  title: "RData Anlysis Project"
author: "Cassidie Stevens, Kassidy McDonald, and Maddie Osgood"
date: "Nov 2021"
output:
  html_document:
  theme: cerulean
highlight: pygments
toc: true
toc_float:
  collapsed: true
smooth_scroll: false
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(tidycensus)
library(janitor)
library(lubridate)
library(sf)
library(tigris)
install.packages("usmap")
library(usmap)

```

```{r}
install.packages("censusxy")
library(censusxy)
```

```{r}
Shootings<-read_csv("data_analysis/data/fatal-police-shootings-data.csv")

census_api_key("0105753aba54d204fb6c33c44044ae59449b3773", overwrite =TRUE)
```
###Question 1:
Are Police shootings primarily only happening in majority Black neighborhoods even when the victim is not black?
###Answer 1:
Yes, it is happening more predominatley in black neighborhoods, the percentage overall in the US for the black population is 12%, and the shooting data showed it as 14% in black areas.
```{r}
shootings<-read_csv("data_analysis/data/fatal-police-shootings-data.csv")%>%
    filter(!is.na(longitude)) 

shootings_geocoded <- tibble()


for (row_number in 1:nrow(shootings_geocoded)) {

  #making one row dataframe with shooting data geocoded
  row_df <- shootings %>%
    slice(row_number)
  #extract value for each object latitude and longitude
  longitude <- row_df$longitude
  latitude <- row_df$latitude
  #get census tract for row
  census_results <- cxy_geography(longitude, latitude) %>%
    select(Census.Tracts.GEOID) %>%
    clean_names()
#add census tract to dataframe 
   row_df <- row_df %>%
     bind_cols(census_results) 
   
  shootings_geocoded <- shootings_geocoded %>%
    bind_rows(row_df) 
}
```
```{r}
#read in the data
#data1 <- read_rds("data_analysis/data/geocoded_results_500.rds")
#data2 <- read_rds("data_analysis/data/geocoded_results_1000.rds")
#data3 <- read_rds("data_analysis/data/geocoded_results_1500.rds")
#data4 <- read_rds("data_analysis/data/geocoded_results_2000.rds")
#data5 <- read_rds("data_analysis/data/geocoded_results_2500.rds")
#data6 <- read_rds("data_analysis/data/geocoded_results_3000.rds")
#data7 <- read_rds("data_analysis/data/geocoded_results_3500.rds")
#data8 <- read_rds("data_analysis/data/geocoded_results_4000.rds")
#data9 <- read_rds("data_analysis/data/geocoded_results_4500.rds")
#data10 <- read_rds("data_analysis/data/geocoded_results_5000.rds")
#data11 <- read_rds("data_analysis/data/geocoded_results_5500.rds")
#data12 <- read_rds("data_analysis/data/geocoded_results_6000.rds")
#data13 <- read_rds("data_analysis/data/geocoded_results_6150.rds")
#data14 <- read_rds("data_analysis/data/geocoded_results_6240.rds")
#data15 <- read_rds("data_analysis/data/geocoded_results_6290.rds")
#data16 <- read_rds("data_analysis/data/geocoded_results_6300.rds")
#data17 <- read_rds("data_analysis/data/geocoded_results_6310.rds")
#data18 <- read_rds("data_analysis/data/geocoded_results_6320.rds")
#data19 <- read_rds("data_analysis/data/geocoded_results_6330.rds")
#data20 <- read_rds("data_analysis/data/geocoded_results_6340.rds")
#data21 <- read_rds("data_analysis/data/geocoded_results_6350.rds")
#data22 <- read_rds("data_analysis/data/geocoded_results_6360.rds")
#data23 <- read_rds("data_analysis/data/geocoded_results_6370.rds")
#data24 <- read_rds("data_analysis/data/geocoded_results_6380.rds")
#data25 <- read_rds("data_analysis/data/geocoded_results_6390.rds")
#data26 <- read_rds("data_analysis/data/geocoded_results_6400.rds")
#data27 <- read_rds("data_analysis/data/geocoded_results_6410.rds")
#binding the ones we have into one
#shootings_geoid <- bind_rows(data1, data2, data3, data4, data5, data6, data7, data8, data9, data10, data11, data12, data13,data14, data15,data16, data17, data18, data19, data20, data21, data22, data23, data24, data25, data26, data27,.id = NULL)

#FILE WE WORK WITH FROM NOW ON
shootings_geoid<- read_rds("data_analysis/all_geocoded_shootings.rds")

unmatched <- anti_join(Shootings, shootings_geoid, by=c("id"))

states_dataframe <- fips_codes %>%
  distinct(state) %>%
  head(51) %>%
  as_vector()

census <- get_acs(geography = "tract", state= states_dataframe, 
                       variables = c(popululation = "B02001_001",
                                median.gross.rent = "B25064_001",
                                median.household.income = "B19013_001",
                                rent.burden = "B25071_001",
                                white = "B03002_003", 
                                af.am = "B03002_004",
                                hispanic = "B03002_012",
                                am.ind = "B03002_005",
                                asian = "B03002_006",
                                nh.pi = "B03002_007",
                                 multiple = "B03002_009",
                                other = "B03002_008"),
                          year= 2019)

write_rds(census,"census.rds")
census <- read_rds("census.rds")

shootings_geoid_grouped <- shootings_geoid%>%
  group_by(census_tracts_geoid)%>%
  summarise(count=n())

census_1 <- census %>%
  select(-moe) %>%
  pivot_wider(names_from=variable, values_from=estimate)

census_added<-census_1%>%
  summarise(total_population = sum(popululation),
            total_black_population = sum(af.am))%>%
mutate(percentage_black = (total_black_population/total_population))

shootings_geoid_rename<-shootings_geoid_grouped%>%
  rename(GEOID=census_tracts_geoid)

shootings_added <- shootings_geoid_rename %>%
  left_join(census_1, by=c("GEOID")) 

shootings_added %>%
  summarise(total_black = sum(af.am),
           total_pop = sum(popululation)) %>%
  mutate(percentage_black=(total_black/total_pop))

```
###Question 2:
Do these shootings mostly occur in low income areas?

###Answer2:
Yes, shootings are occurring more in low income areas. The mean income in the shooting data is $56,554, where the national mean income is $67,007.
```{r}
shootings_income<-shootings_added%>%
  filter(!is.na(median.household.income)) %>%
  summarise(mean_income =mean(median.household.income))

census_income<-census_1%>%
  filter(!is.na(median.household.income)) %>%
  summarise(mean_income =mean(median.household.income))


```

###Question 3:
Are police shootings still happening disproportionately affecting black people regardless of the area (even in white neighborhoods)?

###Answer 3:
Yes, police shootings are disproportionately effect Black people. The rate they get shot is at 23%, it is almost twice as much as there are Black people in the country at 12%

```{r}
options(scipen = 999)

race_shootings <- shootings_geoid %>%
  filter(!is.na(race))

race_counted<-race_shootings%>%
 group_by(race)%>%
  count(race)

race<-race_counted%>%
  add_column(total="6410")

race$total <- as.numeric(race$total)

race_percent<-race%>%
   mutate(percent=(n/total))
```

###PLace to go to tell the story
###Answer:

```{r}
shootings_geoid_black<-shootings_geoid%>%
  filter(race =="B")

shootings_geoid_black<-shootings_geoid_black%>%
  group_by(state)%>%
  summarise(count=n())%>%
  arrange(desc(count))

shooting_state <- shootings_geoid%>%
    group_by(state)%>%
  summarise(count=n())%>%
  arrange(desc(count))%>%
  rename(total_count = count)

shootings_black_state <- shooting_state %>%
  left_join(shootings_geoid_black, by=c("state")) 

shootings_state_percent<-shootings_black_state %>%
   mutate(percentage_black_shot=(count/total_count))%>%
  arrange(desc(percentage_black_shot))

census_2 <- get_acs(geography = "state",
                       variables = c(popululation = "B02001_001",
                                median.gross.rent = "B25064_001",
                                median.household.income = "B19013_001",
                                af.am = "B03002_004"),
                          year= 2019)

write_rds(census_2,"census_2.rds")
census_2 <- read_rds("census_2.rds")

census_3 <- census_2 %>%
  select(-moe) %>%
  pivot_wider(names_from=variable, values_from=estimate)%>%
 mutate(percentage_black=(af.am/popululation))%>%
  arrange(desc(percentage_black))%>%
  rename(state_name = NAME)


fips_codes <- fips_codes %>%
  select(state,state_name) %>%
  distinct()

shootings_black_joined <- fips_codes %>%
  left_join(shootings_state_percent, by=c("state")) %>%
  left_join(census_3, by=c("state_name"))%>%
  mutate(difference=(percentage_black_shot-percentage_black))%>%
  arrange(desc(difference))


```

