---
title: "lab_01"
author: "sean mussenden"
date: "8/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**). 
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**). 
* write out the answer in the form of a complete sentence in the space given (look for **A**). 
When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS. 

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse.To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}


# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
```

## Load Data 

We're going to load the slice of PPP loans for West Virginia.

**Task** Run the codeblock below to load the data. 

```{r}
wv_ppp_loans <- read_rds("lab_01.rds")
```

## Examine Data

**Task** Run the codeblock below to use glimpse to get a sense of the column names, types and sample data. 
**Q1** How many rows and columns does this dataframe have?
**A1** [write your answer here]
33,703 rows and 63 columns.
```{r}
glimpse(wv_ppp_loans)
```

## Answer questions

**Q2** Which West Virginia county got the most loans? How many?
**A2** Kanawha county with 4332 loans.

```{r}
 # Write the code to produce your answer here
 wv_ppp_loans %>%
  group_by(project_county_name) %>%
  summarise(
    count_loans = n()
  ) %>%
  arrange(desc(count_loans))

```
**Q3** Which West Virginia county got the fewest loans? How many?
**A3** Wirt county with 36 loans.
```{r}
# Write the code to produce your answer here
 wv_ppp_loans %>%
  group_by(project_county_name) %>%
  summarise(
    count_loans = n()
  ) %>%
  arrange(count_loans)
```
**Q4** What's your best educated guess to explain why the county identified in question 2 had a lot of loans and why the county identified in question 3 had only a few loans? Hint: you won't find the answer in the data, you'll need to do some quick web research. You do not need to write any code for this question, so no codeblock is provided.  
**A4** 
I think that wirt county had the fewest amount of loans because it has the smallest population of 5717 residents, and Kanawha had the largest number of loans given out because it has the biggest population.

**Q5** What West Virginia business got the state's largest PPP loan? How much was it for? What does the business do? How many jobs did the loan retain, according to the data?
**A5** 
Greenbriar Hotel Corp. had the largest ppp loan of $8,912,202. It retained 500 jobs and the company is based in Sulphur Springs WV, and is apart of the Traveler Acommodation Industry with 3 locations.
```{r}
# Write the code to produce your answer here
 wv_ppp_loans %>%
  arrange(desc(amount))
```

**Q6** What West Virginia politician (with his family) owns the company identified in question 2? What's your source for that information? Hint: this will require web research, you won't be able to answer this purely from the data set.  You do not need to write any code for this question, so no codeblock is provided.  
**A6** James C Justice III is the CEO. https://www.dnb.com/business-directory/company-profiles.greenbrier_hotel_corporation.e1db40121658957db9cea32db34fe5d4.html#contact-anchor
**Q7**. Produce a summary table that has the following information for West Virginia as a whole (the whole state, not by county):
* count of loans
* total amount of all loans
* the average loan amount
* the median loan amount
* the smallest loan amount
* the largest loan amount

Store it as a variable object called `wv_summary` so that you can examine it with the data viewer. What newsworthy nuggets jump out at you that you could imagine dropping into a news story? Write a few summary sentences with overall findings. 

**A7** Something that stood out as newsworthy was the extreme difference in loan amount. The highest ppp loan given out in West Virgina was $8,912,202 where the smallest was $165. This is newsworthy to see how the money was given out based on the company. 
```{r}
wv_summary <- wv_ppp_loans %>%
  group_by(state) %>%
  summarise(
    count_loans = n(),
    total_loan_amount =
  sum(amount),
    average_loan =mean(amount),
    median_loan=
  median(amount),
    smallest_loan = min(amount),
    largest_loan = max(amount)
  )
```

**Q8** Produce a summary table that has the following information for each West Virginia county:
* count of loans
* total amount of all loans
* the average loan amount
* the median loan amount
* the smallest loan amount
* the largest loan amount

Store it as a variable object called `wv_county_summary` so that you can examine it with the data viewer. 

What newsworthy nuggets or patterns jump out at you that you could imagine dropping into a news story or doing follow up reporting on? Write up a few sentence that summarizes one county. 

**A8** A story based on this data that I think would be good to do following up on is how the county with the most loans used them. Kanawha county gave out the most loans of 4332. It would be good to go back and check in on how the money is being used since they got the most amount of loans. 
```{r}
wv_summary <- wv_ppp_loans %>%
  group_by(project_county_name) %>%
  summarise(
    count_loans = n(),
    total_loan_amount =
  sum(amount),
    average_loan =mean(amount),
    median_loan=
  median(amount),
    smallest_loan = min(amount),
    largest_loan = max(amount)
  )
```
